Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-11-27 paper/paper.tex
Christine's comments

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which \emph{preserves adjacency}---that is, if it maps adjacent vertices to adjacent vertices.
The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent
vertices, preserving non-adjacency as well. When working with labelled graphs, a subgraph
isomorphism must preserve labels, and on directed graphs, it must preserve orientation. A
\emph{common induced subgraph} of two graphs $G$ and $H$ is a pair of induced subgraph isomorphisms
from a pattern graph $P$, one to $G$ and one to $H$. A \emph{maximum common induced subgraph} is one
with as many vertices as possible. (The \emph{maximum common partial subgraph} problem is
non-induced, with as many edges as possible; this paper discusses only induced problems.) A common
variant of the problem requires a largest \emph{connected} subgraph
\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.

@@

\begin{description}
    \item[Unlabelled] instances, by selecting the first ten members of each parameter class where the
        graphs have up to 50 vertices each---this gives us a total of 4,110 instances.

    \item[Vertex labelled] instances, by selecting the first ten members of each parameter class
        (and so graphs have up to 100 vertices each), applying the 33\% labelling scheme for
        vertices only, and treating edges as undirected. This gives 8,140 instances.

@@

    \item[Unlabelled, connected] instances, using the same instances as the \emph{unlabelled} case.

@@

\begin{figure}[p]
    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Both labelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize directed}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.3 of C2.east] (C3) {\includegraphics*{gen-graph-pp-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.2 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
    \end{tikzpicture}

@@

In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
profile is a cumulative plot of the relative performance of the parallel algorithm to the sequential
algorithm---instances occurring to the left of the vertical $x = 1$ line are speedups. Each plot
shows three options as different lines. The `all' line include easy instances whose sequential
runtime is below 500 milliseconds, whilst the other two lines include them. The `hard' line treats
sequential timeouts as having been solved at the time limit, whilst the `PAR10' line treats timeouts
as taking ten times longer than the timeout (this convention is common in portfolios
\cite{DBLP:conf/aaai/XuHL10}).

Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
instances on the top row, if we include easy instances, it appears that parallelism is better on
just over half of the instances, whilst if we exclude them, parallelism is always beneficial. If we
do not use the PAR10 scheme, the performance profile also suggests that there are around 500
instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances are
rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
region of 10 (and this is a particular inconvenience because we seek a way of characterising
speedups which does not rely upon us already knowing that 10 is a reasonable choice of penalty).

@@

What about scalability and reproducibility? The first plot in \cref{figure:cliquescale} show the
effects of going from sequential to threaded with two cores, and the next four plots show the
effects of doubling the number of threads each time. These plots show that most of the superlinear
effects occur with fairly small numbers of threads, with nearly all of the benefits of increased
diversity in search being obtained once eight threads are used. As expected, in no case does
increasing the number of threads make things substantially worse.
The final plot in \cref{figure:cliquescale} shows that runtimes are reproducible: running the same
instance on the same hardware twice takes almost exactly the same amount of time.

>>>>

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which \emph{preserves adjacency}---that is, it maps adjacent vertices to adjacent vertices.
The isomorphism is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent
vertices, preserving non-adjacency as well. When working with labelled graphs, a subgraph
isomorphism must preserve labels, and on directed graphs, it must preserve orientation. A
\emph{common induced subgraph} of two graphs $G$ and $H$ is a pair of induced subgraph isomorphisms
from a pattern graph $P$, one to $G$ and one to $H$. A \emph{maximum common induced subgraph} is one
with as many vertices as possible. (The \emph{maximum common partial subgraph} problem is
non-induced, with as many edges as possible; this paper discusses only induced problems.) A common
variant of the problem requires a largest \emph{connected} subgraph
\cite{DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.

@@

\begin{description}
    \item[Unlabelled] undirected instances, by selecting the first ten members of each parameter
        class where the graphs have up to 50 vertices each---this gives us a total of 4,110
        instances.

    \item[Vertex labelled] undirected instances, by selecting the first ten members of each
        parameter class (and so graphs have up to 100 vertices each), and applying the 33\%
        labelling scheme for vertices only. This gives 8,140 instances.

@@

    \item[Unlabelled, connected] instances, as per the \emph{unlabelled} case.

@@

\begin{figure}[p]
    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-plain-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33v-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33ved-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Both labelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize directed}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-plain-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Unlabelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
    \end{tikzpicture}

    \centering
    \begin{tikzpicture}%{{{
        \node[anchor=west] (C1) {\includegraphics*{gen-graph-scatter-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C1.east] (C2) {\includegraphics*{gen-graph-histogram-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, right = 0.1 of C2.east] (C3) {\includegraphics*{gen-graph-pp2-33ve-connected-clique-vs-clique-par-t32.pdf}};
        \node[anchor=west, left = 0.1 of C1.east] (C0) at (0, 0) {\rotatebox{90}{\scriptsize Vertex labelled,}};
        \node[anchor=west, right = -0.2 of C0.east] (C0) {\rotatebox{90}{\scriptsize connected}};
    \end{tikzpicture}

@@

In the third column we show performance profiles \cite{DBLP:journals/mp/DolanM02}. A performance
profile is a cumulative plot of how many times worse the performance of an algorithm is relative to
the virtual best algorithm. Each plot shows three options as different lines. The `all' lines
include easy instances whose sequential runtime is below 500 milliseconds, whilst the other two
lines exclude them. The `hard' line treats sequential timeouts as having been solved at the time
limit, whilst the `PAR10' line treats timeouts as taking ten times longer than the timeout (this
convention is common in portfolios \cite{DBLP:conf/aaai/XuHL10}). The solid lines show the
sequential algorithms, whilst the dotted lines show the parallel algorithms. (There are no dotted
lines on the top four plots for the `hard' and `PAR10' cases, since the parallel algorithm always
beats the sequential algorithm in these cases.) We have normalised the $y$-axis to the number of
counted instances in a given class.

Unfortunately, these three lines can paint very different pictures. For example, for unlabelled
instances on the top row, if we include easy instances, it appears that the parallel algorithm can
be up to ten times worse, whereas if we exclude them, it is never worse. If we do not use the PAR10
scheme, the performance profile also suggests that there are around twenty-five percent of the
hard instances where the speedup is below 10, whilst using PAR10 correctly shows that such instances
are rare. However, PAR10 is only effective in this regard because the ``typical'' speedup is in the
region of 10 (and this is a particular inconvenience because we seek a way of characterising
speedups which does not rely upon us already knowing that 10 is a reasonable choice of penalty).

@@

What about scalability and reproducibility? The first plot in \cref{figure:cliquescale} shows the
effects of going from sequential to threaded with two cores, and the next four plots show the
effects of doubling the number of threads each time. These plots show that most of the superlinear
effects occur with fairly small numbers of threads, with nearly all of the benefits of increased
diversity in search being obtained once eight threads are used. As expected, in no case does
increasing the number of threads make things substantially worse.
The final plot in \cref{figure:cliquescale} shows that runtimes are reproducible: running the same
instance on the same hardware twice takes almost exactly the same amount of time.
